\addcontentsline{toc}{section}{Acknowledgement}
\renewcommand{\abstractname}{Acknowledgement}
\begin{abstract}
First of all, I would like to thank my supervisor J. J. Collins for introducing me this challenging project and providing excellent guidance and support throughout. Without these support and guidance this project would have been nearly impossible to show any substantial outcome.

I want to thank Dr. Norah Power for giving me the opportunity to study this course in CSIS department. I remember the day when I submitted my application for admission still unsure whether I would be accepted or not. And the day when Dr. Power met us in this department for a introduction and a confirmed of a place.

I also want to thank Mr. Tom Mulcahy of Intel Shannon for believing in me and awarding me the \textbf{\textit{Intel Shannon Paul Whelan Scholarship}}. Having received this scholarship meant a lot. It kept my motivation strong throughout.

Many thanks to my family and friends that supported me doing this difficult course for four long years as a mature student. It was a very long journey but I know it is worth every single moment that I spent in this University. I hope you all can forgive me for being so distant sometimes. I hope I can make it up to you all.

Thanks to Josh, Enda, Ben, Angie, Seun, David, Ankit just to name a few that I have had the pleasure to work with as a team on many projects.

Last but not least, a BIG thanks to all CSIS academics and staff members. \newline \newline

\textit{All my works are dedicated to my two little stars; Lamya and Ayaan.}

\end{abstract}